---
title: "4. Benchmarks of lasR vs. lidR"
author: "Jean-Romain Roussel"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{4. Benchmarks of lasR vs. lidR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align='center', fig.width=6, fig.height=4)
col1 = rgb(0, 0.5, 1, alpha = 0.5)
col2 = rgb(1, 0, 0, alpha = 0.5)
col3 = rgb(0.5, 0.5, 1, alpha = 0.5)
col4 = rgb(1, 0, 1, alpha = 0.5)
```

This vignette presents benchmarks for various tasks using `lidR` and `lasR`. The x-axis represents the time spent on the task, while the y-axis represents the memory used for the task.

- **Number of files:** 4
- **Number of points:** 119 million (30 million per file)
- **Coverage:** 10.3 km² (2.5 km² per file)
- **Density:** 12 points/m²
- **OS:** Linux
- **CPU:** 
  - Intel Core i7-5600U CPU (5th generation Intel Core) 
  - Intel Core i7-1355U (13th generation Intel Core) 
- **Date:** update 2024-02-07
- **Version:** 
  - lidR 4.1.1
  - lasR 0.1.0

## Canopy Height Model


```{r, echo=FALSE}
f = c("/home/jr/Documents/Ulaval/ALS data/BCTS/092L073113_BCTS_2.laz", 
"/home/jr/Documents/Ulaval/ALS data/BCTS/092L073114_BCTS_2.laz", 
"/home/jr/Documents/Ulaval/ALS data/BCTS/092L073131_BCTS_2.laz", 
"/home/jr/Documents/Ulaval/ALS data/BCTS/092L073132_BCTS_2.laz")
```

### lidR

```r
chm = rasterize_canopy(ctg, 1, p2r())
```

### lasR

```r
pipeline = reader(ctg) + rasterize(1, "max")
processor(pipeline)
```

### Result

```{r, warning=F, echo=-1, echo=FALSE, fig.show="hold", out.width="50%"}
# 1 
m_lasr = c(0, 93.38, 93.38, 93.38, 93.38, 93.38, 93.38, 93.38, 122.01, 122.01, 122.01, 122.01, 122.01, 122.01, 122.01, 122.53, 122.53, 122.53, 122.53, 122.53, 122.53, 141.60, 
141.60, 141.60, 141.60, 141.60, 141.60, 141.60, 140.92, 149.46,0)
t_lasr = (1:length(m_lasr)-1)*2/60

m_lidr = c(0,104.70, 342.79, 571.07, 793.16, 1028.55, 1388.56, 2027.17, 1888.13, 2164.38, 2381.46, 2519.50, 2280.10, 2519.86, 2783.66, 2984.50, 2266.93, 2379.23, 2635.35, 
2868.93, 2997.68, 2127.28, 0)
t_lidr = (1:length(m_lidr)-1)*5/60

m_lidr2 = c(0, 193.67, 351.17, 499.30, 648.55, 783.17, 931.17, 1079.79, 1132.39, 1197.64, 1972.39, 2222.19, 2372.32, 2521.19, 2690.99, 2839.99, 2946.35, 3068.06, 2392.09, 2670.97, 2818.85, 2967.72, 3137.86, 3293.99, 3370.11, 2619.88, 2307.91, 1989.16, 2128.16, 2286.00, 2433.63, 2579.13, 2656.12, 2723.62, 2594.85, 0)
t_lidr2 = (1:length(m_lidr2)-1)*2/60

m_lasr2 = c(0,203.04, 358.92, 358.92, 371.79, 371.79, 371.79, 372.04, 372.04, 391.29, 391.29, 391.29, 0)
t_lasr2 = (1:length(m_lasr2)-1)*4/60

tmax = max(t_lidr, t_lasr, t_lidr2, t_lasr2)
mmax = max(m_lidr, m_lasr, m_lidr2, m_lasr2)/1000

# Thinkpad

x = t_lidr
y = m_lidr/1000

par(mar = c(4,4,1,1))

plot(x, y, type = "n", xlim = c(0, tmax), ylim = c(0, mmax), ylab = "Memory (GB)", xlab = "Time (min)", main = "Intel Core i7-5600U")

polygon(c(x, rev(x)), c(rep(0, length(x)), rev(y)), col = col1, border = NA)
lines(x, y, col = "blue", lwd = 2)

x = t_lasr
y = m_lasr/1000

polygon(c(x, rev(x)), c(rep(0, length(x)), rev(y)), col = col2 , border = NA)
lines(x, y, col = "red", lwd = 2)

legend("topleft", legend = c("lidR 4.1.1", "lasR 0.1.0"), fill = c(col1, col2), border = NA)

# Latitude

x = t_lidr2
y = m_lidr2/1000

plot(x, y, type = "n", xlim = c(0, tmax), ylim = c(0, mmax), ylab = "Memory (GB)", xlab = "Time (min)", main = "Intel Core i7-1355U")

polygon(c(x, rev(x)), c(rep(0, length(x)), rev(y)), col = col1 , border = NA)
lines(x, y, col = "blue", lwd = 2)

x = t_lasr2
y = m_lasr2/1000

polygon(c(x, rev(x)), c(rep(0, length(x)), rev(y)), col = col2 , border = NA)
lines(x, y, col = "red", lwd = 2)


legend("topleft", legend = c("lidR 4.1.1", "lasR 0.1.0"), fill = c(col1, col2), border = NA)
```

## Digital Terrain Model

```{r, echo=FALSE}
f = c("/home/jr/Documents/Ulaval/ALS data/BCTS/092L073113_BCTS_2.laz", 
"/home/jr/Documents/Ulaval/ALS data/BCTS/092L073114_BCTS_2.laz", 
"/home/jr/Documents/Ulaval/ALS data/BCTS/092L073131_BCTS_2.laz", 
"/home/jr/Documents/Ulaval/ALS data/BCTS/092L073132_BCTS_2.laz")
```

### lidR

```r
dtm = rasterize_terrain(ctg, 1, tin())
```

### lasR

```r
tri = triangulate(filter = keep_ground())
pipeline = reader(ctg, filter = keep_ground()) + tri + rasterize(1, tri)
processor(pipeline)
```

### Result

```{r, warning=F, echo=-1, echo=FALSE, fig.show="hold", out.width="50%"}
# 2
m_lasr = c(0, 79.67, 93.66, 116.75, 133.85, 146.74, 165.05, 257.17, 455.17, 767.38, 833.30, 833.30, 833.30, 846.70, 854.83, 862.04, 869.00, 874.93, 874.93, 894.53, 853.06, 853.06, 
863.89, 872.39, 889.12, 902.23, 618.84, 1002.46, 1013.65, 1013.65, 1022.16, 1041.49, 1049.07, 1055.26, 1061.45, 1061.45, 177.75, 0)
t_lasr = (1:length(m_lasr)-1)*3/60

m_lidr = c(0,124.56, 356.59, 376.44, 966.76, 2048.83, 1790.97, 1790.97, 1754.86, 1754.86, 2457.66, 2169.17, 2169.17, 2170.71, 2304.78, 2654.14, 2172.81, 2195.75, 2170.75, 2231.33, 2579.38, 0)
t_lidr = (1:length(m_lidr)-1)*10/60

m_lidr2 = c(0, 139.30, 346.42, 356.80, 366.42, 372.17, 381.80, 387.30, 414.42, 748.34, 809.84, 890.09, 970.47, 1050.59, 1130.22, 1209.59, 1289.47, 1271.14, 1271.14, 781.42, 787.67, 794.42, 803.29, 811.79, 819.67, 765.35, 925.04, 1001.29, 1076.16, 1151.04, 1225.79, 1300.66, 1375.54, 1450.16, 1392.50, 1392.50, 896.18, 905.18, 919.43, 934.05, 940.18, 923.84, 1020.86, 1101.36, 1181.48, 1260.86, 1339.61, 1418.61, 1497.61, 1576.73, 1656.36, 1728.61, 1651.25, 1651.25, 1030.53, 1039.41, 1047.53, 1054.41, 1060.03, 1063.78, 1101.87, 1195.12, 1270.50, 1345.12, 1420.37, 1495.37, 1570.50, 1645.12, 1601.39, 1601.39, 0)
t_lidr2 = (1:length(m_lidr2)-1)*2/60

m_lasr2 = c(0,202.92, 349.05, 355.55, 372.67, 389.55, 410.09, 422.72, 755.83, 1118.58, 1082.29, 1082.29, 1082.41, 1101.29, 1111.29, 1120.04, 1123.91, 1143.54, 1102.06, 1102.06, 1119.81, 1137.43, 1150.81, 891.28, 1312.16, 1262.08, 1262.08, 1273.33, 1282.21, 1289.33, 1295.33, 1300.96, 1318.83, 0)
t_lasr2 = (1:length(m_lasr2)-1)*2/60

tmax = max(t_lidr, t_lasr, t_lidr2, t_lasr2)
mmax = max(m_lidr, m_lasr, m_lidr2, m_lasr2)/1000

# Thinkpad

x = t_lidr
y = m_lidr/1000

par(mar = c(4,4,1,1))

plot(x, y, type = "n", xlim = c(0, tmax), ylim = c(0, mmax), ylab = "Memory (GB)", xlab = "Time (min)", main = "Intel Core i7-5600U")

polygon(c(x, rev(x)), c(rep(0, length(x)), rev(y)), col = col1, border = NA)
lines(x, y, col = "blue", lwd = 2)

x = t_lasr
y = m_lasr/1000

polygon(c(x, rev(x)), c(rep(0, length(x)), rev(y)), col = col2 , border = NA)
lines(x, y, col = "red", lwd = 2)

legend("topleft", legend = c("lidR 4.1.1", "lasR 0.1.0"), fill = c(col1, col2), border = NA)

# Latitude

x = t_lidr2
y = m_lidr2/1000

plot(x, y, type = "n", xlim = c(0, tmax), ylim = c(0, mmax), ylab = "Memory (GB)", xlab = "Time (min)", main = "Intel Core i7-1355U")

polygon(c(x, rev(x)), c(rep(0, length(x)), rev(y)), col = col1 , border = NA)
lines(x, y, col = "blue", lwd = 2)

x = t_lasr2
y = m_lasr2/1000

polygon(c(x, rev(x)), c(rep(0, length(x)), rev(y)), col = col2 , border = NA)
lines(x, y, col = "red", lwd = 2)


legend("topleft", legend = c("lidR 4.1.1", "lasR 0.1.0"), fill = c(col1, col2), border = NA)
```

## Multiple raster

The gain in terms of computation time is much more significant when running multiple stages in a single pipeline because files are read only once in `lasR` but multiple times in `lidR`. Here, all operations are executed in a single pass at the C++ level, resulting in more efficient memory management.

### lidR

```r
custom_function = function(z,i) { list(avgz = mean(z), avgi = mean(i)) }
ctg = readLAScatalog(f)
chm = rasterize_canopy(ctg, 1, p2r())
met = pixel_metrics(ctg, ~custom_function(Z, Intensity), 20)
den = rasterize_density(ctg, 5)
```

### lasR

```r
custom_function = function(z,i) { list(avgz = mean(z), avgi = mean(i)) }
read = reader(folder)
chm = rasterize(1, "max")
met = rasterize(20, custom_function(Z, Intensity))
den = rasterize(5, "count")
pipeline = read + chm + met + den
processor(pipeline)
```



```{r, warning=F, echo=-1, echo=FALSE, fig.show="hold", out.width="50%"}
# 3
m_lasr = c(0, 82.03, 170.51, 354.01, 531.90, 711.60, 895.17, 1074.86, 1214.60, 1259.20, 1297.35, 1297.35, 435.67,550.40, 692.43, 831.39, 971.39, 1113.44, 1282.57, 1285.92, 1345.21, 
1345.21, 498.85, 559.43, 697.85, 837.33, 977.29, 1127.08, 1258.82, 1240.92, 1240.92, 1240.92, 480.92, 606.07, 749.79, 887.94, 1026.13, 1169.48, 1358.71, 1378.30, 1423.42, 1423.42, 
554.91, 560.07, 0)
t_lasr = 1:length(m_lasr)*2/60

m_lidr = c(0, 104.86, 456.99, 802.07, 1116.71, 1765.46, 2116.21, 2502.28, 2276.53, 2666.14, 2986.86, 2300.90, 2686.19, 2989.73, 2004.21, 2579.39, 3195.28, 2827.48, 3512.35, 4633.82, 
5221.53, 5639.37, 3357.23, 4434.14, 5110.58, 4302.41, 3457.53, 4052.46, 4702.15, 4430.27, 4585.12, 4937.42, 5243.30,4713.65, 5063.31, 5376.67, 2866.46, 3207.25, 3459.13, 3188.25, 
3523.44, 3733.16, 0)
t_lidr = 1:length(m_lidr)*8/60

m_lidr2 = c(0, 193.23, 497.98, 784.95, 1085.33, 1205.75, 2226.83, 2527.83, 2847.46, 2536.90, 2678.37, 2974.99, 3299.73, 2721.75, 1991.53, 2285.40, 2575.28, 2724.10, 2548.08, 3019.45, 3524.15, 3900.47, 2601.94, 3582.43, 4037.68, 4587.39, 4970.17, 2331.14, 3414.44, 3911.82, 4499.68, 4979.93, 3151.38, 3652.82, 4215.47, 4644.97, 4193.78, 3546.69, 3603.89, 3863.48, 4237.23, 3809.42, 1696.35, 1958.85, 2231.42, 2436.03, 2867.71, 3133.63, 3347.25, 2549.77, 2823.65, 3088.82, 3268.23, 3359.37, 0)
t_lidr2 = 1:length(m_lidr2)*4/60

m_lasr2 = c(0, 203.05, 355.17, 448.05, 625.42, 802.55, 979.14, 1157.14, 1332.77, 1482.14, 1488.77, 1527.14, 668.36, 668.48, 842.36, 981.61, 1120.61, 1260.98, 1400.61, 1515.23, 1551.98, 1577.36, 730.80, 778.80, 915.80, 1055.17, 1195.42, 1341.80, 1490.67, 1472.86, 1472.86, 712.91, 765.78, 904.41, 1044.28, 1180.66, 1318.28, 1453.91, 1598.78, 1633.03, 1633.03, 0)
t_lasr2 = 1:length(m_lasr2)*2/60

tmax = max(t_lidr, t_lasr, t_lidr2, t_lasr2)
mmax = max(m_lidr, m_lasr, m_lidr2, m_lasr2)/1000

x = t_lidr
y = m_lidr/1000

par(mar = c(4,4,1,1))

plot(x, y, type = "n", xlim = c(0, tmax), ylim = c(0, mmax), ylab = "Memory (GB)", xlab = "Time (min)", main = "Intel Core i7-5600U CPU")

polygon(c(x, rev(x)), c(rep(0, length(x)), rev(y)), col = col1, border = NA)
lines(x, y, col = "blue", lwd = 2)

x = t_lasr
y = m_lasr/1000

polygon(c(x, rev(x)), c(rep(0, length(x)), rev(y)), col = col2 , border = NA)
lines(x, y, col = "red", lwd = 2)

legend("topleft", legend = c("lidR 4.1.1", "lasR 0.1.0"), fill = c(col1, col2), border = NA)

# Latitude

x = t_lidr2
y = m_lidr2/1000

plot(x, y, type = "n",  xlim = c(0, tmax), ylim = c(0, mmax), ylab = "Memory (GB)", xlab = "Time (min)", main = "Intel Core i7-1355U")

polygon(c(x, rev(x)), c(rep(0, length(x)), rev(y)), col = col1, border = NA)
lines(x, y, col = "blue", lwd = 2)

x = t_lasr2
y = m_lasr2/1000

polygon(c(x, rev(x)), c(rep(0, length(x)), rev(y)), col = col2 , border = NA)
lines(x, y, col = "red", lwd = 2)

legend("topleft", legend = c("lidR 4.1.1", "lasR 0.1.0"), fill = c(col1, col2), border = NA)
```


## Complex Pipeline

In this complex pipeline, the point cloud is normalized and written to new files. A Digital Terrain Model (DTM) is produced, a Canopy Height Model (CHM) is built, and individual trees are detected. These detected trees are then used as seeds for a region-growing algorithm that segments the trees. The `lasR` pipeline can handle hundreds of laser tiles, while `lidR` may struggle to apply the same pipeline, especially during tree segmentation.


```r
read = reader(ctg)
del = triangulate(filter = "-keep_class 2")
norm = transform_with(del)
dtm = rasterize(1, del)
chm = rasterize(1, "max")
seed = local_maximum(3)
tree = region_growing(chm, seed)
write = write_las()
pipeline = read + del + norm + write + dtm + chm +  seed + tree
ans = processor(pipeline, progress = TRUE)
```

```{r, warning=F, echo=-1, echo=FALSE, fig.show="hold", out.width="50%"}
m_lasr = c(0,191.41, 362.51, 593.35, 896.25, 1194.72, 1818.00, 2323.06, 2323.06, 2323.06, 2323.06, 2323.06, 2323.06, 2323.06, 2323.06, 2323.06, 2323.06, 2323.06, 2247.46, 2122.34,
2192.39, 2192.39, 2192.39, 2197.03, 2197.80, 2198.83, 2199.87,2181.38, 2181.38, 2181.38, 2221.91, 1523.74, 1741.71, 1823.69, 1823.69, 1823.69, 1823.69, 1741.83, 1741.83, 1741.83,
1741.83, 1741.83, 1454.66, 1703.53, 1928.16, 2152.36, 2276.37,2506.08, 2506.08, 2506.08, 2506.08, 2506.08, 2506.08, 2506.08,2506.08, 2506.08, 2506.08, 2506.08, 2277.12, 2324.82,
2324.82, 2324.82, 2324.82, 2324.82, 2324.82, 2324.82, 2324.82, 2296.59, 2296.59, 2296.59, 2323.40, 1519.98, 1738.68, 1900.92, 2009.98, 2009.98, 2009.98, 2009.98, 2009.98, 1901.16,
1901.16, 1901.16, 1901.16, 1901.16, 1901.16, 765.81, 768.27)

m_lidr = c(0,111.37, 370.23, 1302.22, 2092.03, 1836.73, 1777.86, 1778.63, 1772.64, 2176.11, 1783.72, 1788.08, 2751.04, 3687.52, 4444.45, 4444.45, 4444.45, 4444.45, 4444.45, 5891.63, 
6243.98, 5576.12, 5058.68, 6276.28, 5694.89, 5694.89, 5694.89, 5694.89, 4708.56, 6252.17, 7429.84, 5487.51, 5091.87, 4924.78, 2821.48, 4750.88, 4750.88, 4750.88, 4750.88, 4750.88, 
4750.88, 4750.88, 4750.88, 4750.88, 3746.96, 4379.66, 4379.66, 4379.66, 5547.57, 5956.35, 5956.35, 5956.35, 5956.35, 5956.35, 5956.35, 5956.35,5956.35, 5956.35, 4058.14, 4737.40, 
4737.40, 4737.40, 5196.26, 5196.26,0)

t_lasr = (1:length(m_lasr)-1)*5/60
t_lidr = (1:length(m_lidr)-1)*15/60

col1 = rgb(0, 0.5, 1, alpha = 0.5)
col2 = rgb(1, 0, 0, alpha = 0.5)

x = t_lidr
y = m_lidr/1000

par(mar = c(4,4,1,1))

plot(x, y, type = "n", ylim = c(0, max(y)), ylab = "Memory (GB)", xlab = "Time (min)", main = "Intel Core i7-5600U CPU")

polygon(c(x, rev(x)), c(rep(0, length(x)), rev(y)), col = col1, border = NA)
lines(x, y, col = "blue", lwd = 2)

x = t_lasr
y = m_lasr/1000

polygon(c(x, rev(x)), c(rep(0, length(x)), rev(y)), col = col2 , border = NA)
lines(x, y, col = "red", lwd = 2)

legend("topleft", legend = c("lidR 4.1.1", "lasR 0.1.0"), fill = c(col1, col2), border = NA)
```